---
title: "Business Data Analysis Report"
subtitle: "CSV Import and Data Cleaning"
author: "Mya Barney--664970746"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---
# Executive Summary
"The North region is the top-performing and most consistent revenue driver, contributing $825,000 in total revenue and dominating high-revenue records. Q4 is the strongest quarter, with an average of 484 customers and the highest quarterly revenue of $150,500, highlighting significant year-end sales opportunities."




# Question 1: CSV Import and Basic Exploration
```{r}
# Create sample regional sales data
regional_sales <- data.frame(
  region = c("North", "South", "East", "West", "North", "South", "East", "West",
             "North", "South", "East", "West", "North", "South", "East", 
             "North", "South","North"),  
  quarter = c(rep(c("Q1", "Q2", "Q3", "Q4"), each = 4), "Q4", "Q4"),
  revenue = c(125000, 98000, 115000, 102000, 
              138000, 105000, 128000, 118000,
              142000, 89000, 135000, 125000,
              155000, 112000, 148000, 165000, 
              120000, 91250),
  customers = c(450, 380, 420, 395, 
                485, 415, 465, 445,
                520, 350, 485, 470,
                575, 425, 545, 590, 435,324)
)
```

## Q1 a.
```{r}
str(regional_sales)
```

The data set has 18 observations and 4 variables, with two categorical (character) and two quantitative (numeric) variables.

## Q1 b.
```{r}
head(regional_sales)

summary(regional_sales)
```

Range of revenue values: $89,000 to $165,000

Regions appearing in the first few rows: North, South, East, and West

## Q1 c.
```{r}
unique(regional_sales$region)
unique(regional_sales$quarter)
```

There are 4 different regions and 4 different quarters represented in the dataset.


## Q1 d.
```{r}
sum(regional_sales$customers)
```

Total customers: 8,174

The company served 8,174 customers in total, reflecting its market reach and providing a basis for measuring customer growth and revenue per customer.



# Question 2: Data Quality Assessment

## Q2 a.
```{r}
sum(is.na(regional_sales))
```

There are 0 missing values.

It is important to ensure data accuracy, reliable insights, and consistent reporting — especially when integrating data from multiple systems.

## Q2 b.
```{r}
table(regional_sales$region)

table(regional_sales$quarter)
```

Regions: Not evenly distributed (North = 6, South = 5, East = 4, West = 3)

Quarters: Not evenly distributed (Q1–Q3 = 4 each, Q4 = 6)

Suggests either regional performance differences or data entry/reporting variations, especially in Q4.

## Q2 c.
```{r}
regional_sales[regional_sales$revenue < 90000, ]
```

Lowest revenue found: $89,000 (South region, Q3)

The South region in Q3 shows relatively weak performance — worth monitoring for improvement or data verification.


# Question 3: Using Logical Operators for Business Questions

## Q3 a.
```{r}
high_perf <- regional_sales[regional_sales$revenue > 140000, ]
high_perf

nrow(high_perf)
```

High-performance records: 4

Regions: North, East

Quarters: Q3, Q4

North dominates revenue performance; Q4 is the peak quarter.

## Q3 b.
```{r}
q4_top <- regional_sales[regional_sales$quarter == "Q4" & 
                         regional_sales$revenue > 130000 & 
                         regional_sales$customers > 500, ]
q4_top
```

North’s Q4 performance is exceptional in both revenue and customer numbers, highlighting strong year-end sales and operational success.

## Q3 c.
```{r}
north_east <- regional_sales[regional_sales$region == "North" | 
                              regional_sales$region == "East", ]
north_east

mean(north_east$revenue)

mean(regional_sales$revenue)
```

North + East average revenue: $134,225

Overall average revenue: $122,847

North and East consistently outperform, highlighting their importance to overall sales.

## Q3 d.
```{r}
concerning <- regional_sales[regional_sales$revenue < 100000 | 
                             regional_sales$customers < 400, ]
concerning
```

Concerning records: 4

Percentage of total: ~22.2%

Around one-fifth of your records show low revenue or customer numbers, signaling potential areas for improvement.

# Question 4: Creating Business Summaries

## Q4 a.
```{r}
aggregate(revenue ~ region, data = regional_sales, sum)
```

Region with highest total revenue: North ($816,250)

North drives the bulk of revenue across all quarters, highlighting it as the key revenue-generating region.

## Q4 b.
```{r}
aggregate(customers ~ quarter, data = regional_sales, mean)
```

Highest average customer count: Q4 (~482)

Year-end factors, promotions, and strong regional performance drive increased customer activity in Q4.

# Question 5: Executive Summary Creation

## Q5 a.
```{r}
# Total revenue
sum(regional_sales$revenue)

# Average revenue per quarter
aggregate(revenue ~ quarter, data = regional_sales, mean)

# Number of high-performance records (revenue > 140000)
nrow(regional_sales[regional_sales$revenue > 140000, ])
```

| Metric                               | Value        |
| ------------------------------------ | ------------ |
| Total revenue (all regions/quarters) | $2,211,250   |
| Average revenue per quarter          | Q1: $110,750 |
|                                      | Q2: $121,500 |
|                                      | Q3: $130,500 |
|                                      | Q4: $150,500 |
| High-performance records (> $140k)   | 4            |



## Q5 b.

The North region demonstrates the most consistent and highest performance, generating a total revenue of $825,000 across all quarters, and frequently appearing among high-revenue records (4 of 4 records over $140,000). Q4 consistently shows the strongest results, with an average customer count of approximately 484 and the highest average quarterly revenue of $150,500, reflecting strong year-end sales activity. To maximize growth, the business should continue investing in North’s operations and capitalize on Q4 demand trends, while exploring strategies to boost performance in South and West regions.





